---
title: "owner-analysis"
author: "Esha Prasad"
date: "4-16-26"
format: html
editor: visual
---

## Exercise: Explore, Clean, and Wrangle Data

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
```

## Loading the Data

```{r}
lobster_abundance <- read_csv(here::here("data/Lobster_Abundance_All_Years_20220829.csv"))
```

### Exercise 1

```{r}
lobster_abundance <- lobster_abundance %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))
```

This function makes it so that values of "-99999" in the column "SIZE_MM" are turned into "NA", which is easier for R to read when reading through the data table.

### Exercise 2

```{r}
not_napl <- lobster_abundance %>% 
     filter(SITE != "NAPL")
```

This function makes it so that all values from the site "NAPL" are removed in the new table.

### Exercise 3

```{r}
aque_70mm <- lobster_abundance %>% 
     filter(SITE == "AQUE" & SIZE_MM >= 70)
```

This function makes it so that only lobsters from the site "AQUE" over a size of 70mm are put into the new table.

### Exercise 4

```{r}
max_lobster <- lobster_abundance %>% 
     group_by(SITE, MONTH) %>% 
     summarize(MAX_LENGTH = max(SIZE_MM, na.rm = TRUE))
```

This function makes it so that the lobster data is compressed by site and month, and shows the max lobster length at each site for that month.
