---
title: "owner-analysis"
author: "Esha Prasad"
date: "4-16-26"
format: html
editor: visual
---

## Exercise: Explore, Clean, and Wrangle Data

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
```

### Loading the Data

```{r}
lobster_abundance <- read_csv(here::here("data/Lobster_Abundance_All_Years_20220829.csv"))
```

### Exercise 1

```{r}
lobster_abundance <- lobster_abundance %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))
```

This function makes it so that values of "-99999" in the column "SIZE_MM" are turned into "NA", which is easier for R to read when reading through the data table.

### Exercise 2

```{r}
not_napl <- lobster_abundance %>% 
     filter(SITE != "NAPL")
```

This function makes it so that all values from the site "NAPL" are removed in the new table.

### Exercise 3

```{r}
aque_70mm <- lobster_abundance %>% 
     filter(SITE == "AQUE" & SIZE_MM >= 70)
```

This function makes it so that only lobsters from the site "AQUE" over a size of 70mm are put into the new table.

### Exercise 4

```{r}
max_lobster <- lobster_abundance %>% 
     group_by(SITE, MONTH) %>% 
     summarize(MAX_LENGTH = max(SIZE_MM, na.rm = TRUE))
```

This function makes it so that the lobster data is compressed by site and month, and shows the max lobster length at each site for that month.

## Exercise: Data Visualization

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)

lobsters_summarize <- lobster_abundance %>% 
  group_by(SITE, YEAR) %>% 
  summarize(COUNT = n())

lobster_size_lrg <- lobster_abundance %>%
    filter(YEAR %in% c(2019, 2020, 2021)) %>%
    mutate(SIZE_BIN = if_else(SIZE_MM <= 70, true = "small", false = "large")) %>%
    group_by(SITE, SIZE_BIN) %>%
    summarize(COUNT = n()) %>%
    drop_na()
```

### Histogram Creation

```{r}
lobster_histo <- ggplot(lobster_abundance, aes(x=SIZE_MM)) + 
  geom_histogram(fill="blue", color="black", binwidth = 10) + 
  facet_wrap(~SITE) + 
  xlab("Lobster size") + 
  ylab("Number of lobsters") + 
  theme_minimal()

lobster_histo
```

### Line Graph

```{r}
lobster_point <- ggplot(data = lobsters_summarize, aes(x = YEAR, y = COUNT)) +
  geom_point(aes(color = SITE), shape="triangle", size=2) +
  geom_line(aes(color = SITE)) + theme_classic() + 
  xlab("Study Year") + 
  ylab("Number of Lobsters")

lobster_point
```

### Bar Graph

```{r}
lobster_bar <- ggplot(data = lobster_size_lrg, aes(x = SITE, y = COUNT, fill = SIZE_BIN)) +
  geom_col(position = "dodge") + 
  scale_fill_manual("legend", values = c("large" = "darkgreen", "small" = "darkred")) +
  xlab("Site Name") +
  ylab("Number of lobsters")

lobster_bar
```
